---
title: "Nerdle Analysis"
output:
  html_document:
    df_print: paged
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(stringr)
library(dplyr)
library(kableExtra)
source("nerdlefunctions.r")
dfEquations <- readRDS("validEquations.rds")
dfEqForms <- unique(dfEquations$Form)
byFormTbl <- dfEquations %>% group_by(Form) %>% summarize(n = n()) %>% arrange(desc(n))
byEqPos <- byFormTbl %>% mutate(EqPos = stringr::str_locate(byFormTbl$Form, "=")[,1]) %>%
  group_by(EqPos) %>% summarize(Count = sum(n))
parsedNumbers <- t(sapply(dfEquations$Digits, parseDigits, NULL))
temp <- unlist(parsedNumbers)
byDigit <- data.frame(Digit = 0:9, Count = 0)
for (i in 0:9){
  byDigit[i + 1, "Count"] <- sum(temp == i)
}
byOperator <- data.frame(Operator = c("+", "-", "*", "/"), 
                         Count = 0)
for (i in 1:nrow(byFormTbl)) {
  byOperator[byOperator$Operator == "+", "Count"] <-
    byOperator[byOperator$Operator == "+", "Count"] +
    str_count(as.character(unlist(byFormTbl[i, "Form"])),"\\+") *
    as.integer(byFormTbl[i, "n"])
  byOperator[byOperator$Operator == "-", "Count"] <-
    byOperator[byOperator$Operator == "-", "Count"] +
    str_count(as.character(unlist(byFormTbl[i, "Form"])),"-") *
    as.integer(byFormTbl[i, "n"])
  byOperator[byOperator$Operator == "*", "Count"] <-
    byOperator[byOperator$Operator == "*", "Count"] +
    str_count(as.character(unlist(byFormTbl[i, "Form"])),"//*") *
    as.integer(byFormTbl[i, "n"])
  byOperator[byOperator$Operator == "/", "Count"] <-
    byOperator[byOperator$Operator == "/", "Count"] + 
    str_count(as.character(unlist(byFormTbl[i, "Form"])),"\\/") *
    as.integer(byFormTbl[i, "n"])
}

```
Thinking about Nerdle.  I believe all Nerdle equations can be written in one of the following `r length(dfEqForms)` forms.

```{r echo=FALSE}
kable(dfEqForms, caption = "Forms of Nerdle Equations", col.names = "")
```

         
*Have I missed any?*

Note that d + dd - d is the same as d - d + dd so we can arbitrarily pick one.  Leading zeros and single digit zeros are not allowed except that the right hand side may be a single digit zero.

****    

There are **`r nrow(dfEquations)`** valid equations.  Here are the number by form

```{r echo=FALSE}
print(byFormTbl, n = length(dfEqForms))
```

****    

**Position of '=' sign**    
By inspection we see that the '=' sign is most likely to be in the 6th position.  Here are the numbers
```{r echo=FALSE}
print(byEqPos)
```

****    

**Frequencies of Digits**
```{r echo=FALSE}
print(byDigit)
barplot(byDigit$Count, xlab = "Digit", ylab = "Frequency (Count)", names.arg = byDigit$Digit)
```

****    

**Frequencies of Operators**    
```{r echo=FALSE}
print(byOperator)
barplot(byOperator$Count, xlab = "Operator", ylab = "Frequency (Count)", names.arg = byOperator$Operator)
```

Interesting that subtraction is the most common operator.  

****    

**Frequency of digit by location**    
Equations with two operators have five digits (e.g., "d-d+dd=d"). 
Equations with one operator have six digits (e.g., "dd-dd=dd").
Labelling the positions from one to five (or six) from left to right,
we show the frequency of digit by form of equation.

```{r echo=FALSE, results='asis'}
lstDigitsByPosition <- vector("list", length(eqForms))
names(lstDigitsByPosition) <- eqForms
for (eqForm in levels(dfEquations$Form)){
  nDigitsInForm <- stringr::str_count(eqForm, "d")
  dfDigitsByPosition <- data.frame(Digit = 0:9,
                                 Pos1 = 0, Pos2 = 0, Pos3 = 0,
                                 Pos4 = 0, Pos5 = 0, Pos6 = 0)
  idx <- dfEquations$Form == eqForm 
  temp <- data.frame(unlist(t(sapply(dfEquations[idx,]$Digits, parseDigits, NULL))))
  for (d in 0:9){
    tempd <- temp == d
    dByPos <- apply(tempd, 2, sum)
    dfDigitsByPosition[d+1, 2:(nDigitsInForm + 1)] <-
      dfDigitsByPosition[d+1, 2:(nDigitsInForm + 1)] +
      dByPos
  }
  dfDigitsByPosition <- dfDigitsByPosition %>% mutate(Total = Pos1+Pos2+Pos3+Pos4+Pos5+Pos6)
  lstDigitsByPosition[[eqForm]] <- dfDigitsByPosition

  print(kableExtra::kable(dfDigitsByPosition, 
                          caption = paste("Form of Equation:", eqForm), 
                          format = "html"))
}
```

